### 2015-05-22 reports/paper/experiment.tex
paper update. emailed

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_1.JPG}
    \caption{1 meter by 1 meter grid}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_2.JPG}
    \caption{array placement}
  \end{subfigure}
  \caption{Setup for localization accuracy testing}
  \label{fig:setup_point}
\end{figure}

@@

For each experiment conducted, two different movement filters were evaluated:
\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
\item[Averaging filter] localization for past $0.5$ seconds were averaged and outputed as current estimate.
\item[Kalman filter] A 2nd order Kalman filtering was used.
\end{description}

@@

We also calculated the localization error for each tested point in the grid. Figure~\ref{fig:error_distribution} shows the error distribution inside the grid. The error is below $3$ cm for most areas inside the grid. There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike.

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
\caption{accuracy versus window size}
\label{fig:accuracy_vs_window}
\end{figure}

\begin{figure}[]
\centering
\includegraphics[width=0.4\textwidth]{calculation_time.eps}
\caption{speed versus window size}
\label{fig:speed_vs_window}
\end{figure}

@@

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_ring.JPG}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
  \end{subfigure}
  \caption{Setup for circle movement localization}
  \label{fig:setup_circle}
\end{figure}

@@

Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA. Which makes localization noisy. As window size increases, localization converges to the shape of the ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. Error decreases as the window size increases and plateaus after it exceeds around $10$ milliseconds.

@@

Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound sorce is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

>>>>

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_1.JPG}
    \caption{1 meter by 1 meter grid}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_2.JPG}
    \caption{array placement}
  \end{subfigure}
  \caption{Setup for point localization evaluation}
  \label{fig:setup_point}
\end{figure}

@@

For each experiment conducted, two different movement filters were evaluated:
\begin{description}[\IEEEsetlabelwidth{Very long label}\IEEEusemathlabelsep]
\item[Averaging filter] localization for past $0.5$ seconds were averaged and outputted as current estimate.
\item[Kalman filter] A 2nd order Kalman filter was used.
\end{description}

@@

We also calculated the localization error for each tested point in the grid. Figure~\ref{fig:error_distribution} shows the error distribution inside the grid. The error is below $3$ cm for most areas inside the grid. There is one error spike in the mid-left region and we contribute this to audio source misplacement because the error is fairly low and consistent around that spike.

\begin{figure}[]
\includegraphics[width=0.5\textwidth]{average_error_window_size.eps}
\caption{Localization error versus window size}
\label{fig:accuracy_vs_window}
\end{figure}

\begin{figure}[]
\centering
\includegraphics[width=0.4\textwidth]{calculation_time.eps}
\caption{Computation time versus window size}
\label{fig:speed_vs_window}
\end{figure}

@@

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_ring.JPG}
  \end{subfigure}
  \begin{subfigure}[]{.2\textwidth}
    \includegraphics[width=\textwidth]{setup_rotating_disk.JPG}
  \end{subfigure}
  \caption{Setup for movement tracking evaluation}
  \label{fig:setup_circle}
\end{figure}

@@

Fig~\ref{fig:wn} gives an intuitive representation of how accuracy changes with window size. When window size is small($1.02$ millisecond), the audio does not contain enough information to reliably estimate TDOA, which results in noisy localization. As window size increases, localization converges to the shape of the ground truth circle. Fig~\ref{fig:err_win_circle} shows how the error changes with window size. The general trend is similar to that in point localization case. Error decreases as the window size increases and plateaus after it exceeds around $10$ milliseconds.

@@

Fig~\ref{fig:circle_wnn} to \ref{fig:circle_musicbn} shows results for experiments at normal speed, and Fig~\ref{fig:circle_wnf} to~\ref{fig:circle_musicbf} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find that localization error does not depend on how fast the sound source is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

### 2015-05-22 reports/paper/introduction.tex
paper update. emailed

%Localization has been a widely researched topic in both academia and industry due to its wide range of applications. Companies and individuals have bennefited from technologies such as GPS, ultrasound, and blah.  Since xxx  GPS technology has been invented and built into cars for navigation which provided great convenience for drivers. Ultrasound when used by sonar, help greatly helped underwater mapping. However, when it comes to localizationon a smaller scale such as the inside of a room, GPS would not be of much help because it's accuracy range is $1.4$ m~\cite{intro:gps}.

Accurate indoor localization allows creation of novel applications with surrounding awareness that uses position and movement information as input. One application is to allow users to draw with music without physically touching the computer. Another example is to build AI games with physical pieces such as toy car racing where the computer controls some them. In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.

Global Positioning System (GPS) is the prevailing technology used for outdoor localization. Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}. While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking. Ultrasound based indoor localization approaches on the other hand, has achieved sub-centimeter accuracy~\cite{intro:ultra}. However, ultrasound systems require the use of expensive transducers.

@@

In this project, we have built a localization system with reasonable high precision for small area using microphone arrays that localize typical audio source. Our system is built with inexpensive electret microphones mounted on portable frames. Users can interact with our system using any device that has audio output such as a mobile phone.

In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization. In Section III, we evaluated different array architectures and their impact on localization accuracy. We demonstrated that a two array system outperforms a single array system of similar physical dimension. In Section IV, we presented the chosen architecture along with hardware details. Finally, experiment details and results are presented in Section V.

>>>>

Accurate indoor localization allows creation of novel applications with surrounding awareness that uses position and movement information as input. One application is to allow users to draw with music without physically touching the computer. Another example is to build AI games with physical pieces such as toy car racing where the computer controls some of the toy cars. In this work, we aim to build such a source localization system that is portable, inexpensive, yet reasonably accurate for localization in a small area.

Global Positioning System (GPS) is the prevailing technology used for outdoor localization. Commercial grade GPS has an average error of a few meters, depending on the size and quality of the receiver~\cite{intro:gps}. While accuracy in this range is good for many applications including driving navigation and vehicle tracking, it does not provide enough precision for local movement tracking. Ultrasound based indoor localization approaches on the other hand, have achieved sub-centimeter accuracy~\cite{intro:ultra}. However, ultrasound systems require the use of expensive transducers.

@@

In this project, we have built a localization system with reasonably high precision for small area using microphone arrays that localize typical audio sources. Our system is built with inexpensive electret microphones mounted on portable frames. Users can interact with our system using any device that has audio output such as a mobile phone.

In this paper, we first discussed in Section II some prior relevant research and approaches in sound localization. In Section III, we evaluated different array architectures and their impact on localization accuracy. We demonstrated that a two array system outperforms a single array system of similar physical dimension. In Section IV, we presented the chosen architecture along with hardware details. Finally, experiment details and results were presented in Section V.

### 2015-05-22 reports/paper/prior.tex
paper update. emailed

\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user tap on interactive surfaces. \cite{ltm:lpc} used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces. \cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative template.

@@

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
\end{description}

>>>>

\cite{extended:tusi} and \cite{ltm:pham} investigated using max value from cross-correlation as a similarity measure to localize user taps on interactive surfaces. \cite{ltm:lpc} used L2 distance in the Linear Predictive Coding coefficient space as a similarity measure to localize taps on surfaces. \cite{ltm:tusi2} further explored accuracy improvement by using multiple templates for each location and speed improvement by merging multiple templates into one representative template.

@@

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
\end{description}

### 2015-05-22 reports/paper/simulation.tex
paper update. emailed

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_1}
    \caption{source at $(r=50$ cm,$\theta = 0$ degrees$)$}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_2}
    \caption{source at $(r=50$ cm,$\theta = 45$ degrees$)$}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_3}
    \caption{source at $(r=50$ cm,$\theta = 90$ degrees$)$}
  \end{subfigure}
  \caption{Uncertainty region}
  \label{fig:sim_2_5}
\end{figure}

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Let $2a$ denote the difference of distance between $P$ and two microphones:
\[
2a = P M_1 - P M_2 
\]
, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $a$ within $1$ cm:
\[
R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
\]
Intuitively, points in $R$ have difference of distance to two microphones very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertainty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

This can also be seen analytically. Assuming two microphones are placed on the x-axis at $M_1:(-c,0)$ and $M2:(c,0)$. All points $P:(x,y)$ with difference of distance $ |PM_1 - PM_2| = 2a$ satisfies:
\begin{eqnarray}\label{eqn:hyperbola}
\frac{x^2}{a^2} - \frac{y^2}{c^2-a^2} = 1
\end{eqnarray}
To see how difference of distance changes with respect to distance, we can expand the equation and find partial differential $\frac{\partial a}{\partial x}$:
\begin{eqnarray}\label{eqn:derivative}
\frac{\partial a}{\partial x} = \frac{x(c^2-a^2)}{a(x^2+y^2+c^2)-2a^3}
\end{eqnarray}
Since all points in equation~\ref{eqn:derivative} must lie on the hyperbola, we can substitute~\ref{eqn:hyperbola} into~\ref{eqn:derivative}:
\begin{eqnarray}\label{eqn:derivativeF}
\frac{\partial a}{\partial x} = \frac{c^2-a^2}{\frac{c^2}{a}x - \frac{a^3}{x}}
\end{eqnarray}

@@

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{$20$cm equilateral triangle array. Source is $20$cm away from the array}
  \label{fig:sim_3_2}
\end{figure*}

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{$20$ cm equilateral triangle array. Source is $80$ cm away from the array}
  \label{fig:sim_3_8}
\end{figure*}

With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions. The smaller the intersection region, the better the localization accuracy. To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of an $20$ cm equilateral triangle. An audio source is placed at $20$ cm away from the center of the array. Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source. It can be seen that accuracy decreases when sound source becomes close to the line connecting any two microphones. This observation is consistent with the two microphone case, since points close to lines connecting microphones have a larger uncertainty region.

@@

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
    \caption{$3$ microphones placed at vertices of an $20$cm equilateral triangle. Average error is $18.6$ cm}
    \label{fig:sim_hm_3}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
    \caption{another microphone added at origin. Average error is $17.1$ cm}
    \label{fig:sim_hm_3_p1}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
    \caption{$3$ microphones placed at vertices of an $40$cm equilateral triangle. Average error is $10.04$ cm}
    \label{fig:sim_hm_3_2x}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
    \caption{$3$ microphones placed in a line. Total length is $20$ cm. Average error is $55.05$ cm}
    \label{fig:sim_hm_3_line}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
    \caption{$4$ microphones placed at $4$ corners of the region. Average error is $0.05$ cm}
    \label{fig:sim_hm_4}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_2a}
    \caption{two $3$ microphone arrays placed $1$ meter apart. Average error is $2.60$ cm}
    \label{fig:sim_hm_2_array}
  \end{subfigure}
  \caption{Accuracy for different array configurations}
  \label{fig:sim_hm}
\end{figure*}

Each microphone pair generates a hyperbolic region, and the intersection area of these regions is a measure of the localization accuracy. To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid. The center location of intersection region can be used as the localization estimate to calculate localization error. Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.

Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle. The region inside the array has good accuracy. However, for regions along the line connecting any two microphones, the accuracy drops significantly. Average error across the region is $18.6$ cm.

To evaluate how adding one microphone(without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$. Result is presented in fig~\ref{fig:sim_hm_3_p1}. Addition of the new microphone only slightly improved the accuracy around the array region. Average error dropped from $18.6$ cm to $17.1$ cm. Regions near lines connecting microphones still have significantly larger uncertainty region.

To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$. The result is presented in fig~\ref{fig:sim_hm_3_2x}. The overall uncertainty area decreased across the region. Average error improved to $10.04$ cm.

>>>>

\begin{figure}[]
  \centering
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_1}
    \caption{source at $(r=50$ cm, $\theta = 0$ degrees$)$}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_2}
    \caption{source at $(r=50$ cm, $\theta = 45$ degrees$)$}
  \end{subfigure}
  \begin{subfigure}[]{.23\textwidth}
    \includegraphics[width=\textwidth]{sim/sim_2_3}
    \caption{source at $(r=50$ cm, $\theta = 90$ degrees$)$}
  \end{subfigure}
  \caption{Uncertainty region. Yellow dots represent microphones' locations and the white dot represents the location of the source.}
  \label{fig:sim_2_5}
\end{figure}

To see how precision affects localization accuracy, we simulated two microphones placed at: $M_1:(x=-10\mbox{ cm},y=0\mbox{ cm})$ and $M_2:(x=10\mbox{ cm},y=0\mbox{ cm})$. A test sound source is emitted at point $P$ which is $50$ centimeters away from the origin $(0,0)$. Let $2a$ denote the difference of distance between $P$ and two microphones:
\[
2a = P M_1 - P M_2 
\]
, then fig~\ref{fig:sim_2_5} shows the region $R$ where all points have difference of distance to two microphones close to $2a$ within $1$ cm:
\[
R=\{\hat P: |(\hat P M_1 - \hat P M_2) - 2a|< 1 \mbox{ cm}\}
\]
Intuitively, points in $R$ have difference of distance to two microphones very similar to each other. Looking at fig~\ref{fig:sim_2_5}, we can still see that $R$ has the shape of a hyperbola, but with an uncertainty region around it. The thickness of the uncertainty region is not uniform around the hyperbola, the farther away the point is, the larger the uncertainty region becomes. This indicates for the same delta distance movement it will generate smaller difference of distance change when the source is farther away from the array. The size of the uncertainty region is also angle dependent: points closer to the line connecting microphones have larger region compared to points close to the line bisecting microphones.

This can also be seen analytically. Assuming two microphones are placed on the x-axis at $M_1:(-c,0)$ and $M2:(c,0)$. All points $P:(x,y)$ with difference of distance $ |PM_1 - PM_2| = 2a$ satisfies:
\begin{eqnarray}\label{eqn:hyperbola}
\frac{x^2}{a^2} - \frac{y^2}{c^2-a^2} = 1
\end{eqnarray}
To see how the difference of distance changes with respect to source location, we can expand the equation and find the partial differential $\frac{\partial a}{\partial x}$:
\begin{eqnarray}\label{eqn:derivative}
\frac{\partial a}{\partial x} = \frac{x(c^2-a^2)}{a(x^2+y^2+c^2)-2a^3}
\end{eqnarray}
Since all points in equation~\ref{eqn:derivative} must lie on the hyperbola, we can substitute~\ref{eqn:hyperbola} into~\ref{eqn:derivative}:
\begin{eqnarray}\label{eqn:derivativeF}
\frac{\partial a}{\partial x} = \frac{c^2-a^2}{\frac{c^2}{a}x - \frac{a^3}{x}}
\end{eqnarray}

@@

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_20cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $20$cm away from the array.}
  \label{fig:sim_3_2}
\end{figure*}

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_0_degrees}
    \caption{0 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_45_degrees}
    \caption{45 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_90_degrees}
    \caption{90 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_135_degrees}
    \caption{135 degrees}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_80cm_180_degrees}
    \caption{180 degrees}
  \end{subfigure}
  \caption{Uncertainty region. Microphones are at the vertices of a $20$cm equilateral triangle. The source is $80$cm away from the array.}
  \label{fig:sim_3_8}
\end{figure*}

With more than two microphones, each pair of microphones generates a hyperbolic region and localization becomes finding the intersection of hyperbolic regions. The smaller the intersection region, the better the localization accuracy. To see how accuracy changes with array placement and sound source location, three microphones are placed at the three vertices of a $20$ cm equilateral triangle. An audio source is placed at $20$ cm away from the center of the array. Fig~\ref{fig:sim_3_2} shows the intersection of regions for $5$ different placement of the sound source. It can be seen that accuracy decreases when sound source becomes close to the line connecting any two microphones. This observation is consistent with the two microphone case, since points close to lines connecting microphones have a larger uncertainty region.

@@

\begin{figure*}[]
  \centering
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_center}
    \caption{$3$ microphones are placed at vertices of a $20$cm equilateral triangle. The average error is $18.6$ cm}
    \label{fig:sim_hm_3}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_p1}
    \caption{Another microphone is added at origin. The average error is $17.1$ cm}
    \label{fig:sim_hm_3_p1}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_2x}
    \caption{$3$ microphones are placed at vertices of a $40$cm equilateral triangle. The average error is $10.04$ cm}
    \label{fig:sim_hm_3_2x}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_3_line}
    \caption{$3$ microphones are placed in a line, $10$ cm apart from each other. The average error is $55.05$ cm}
    \label{fig:sim_hm_3_line}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_4corner}
    \caption{$4$ microphones are placed at $4$ corners of the grid. The average error is $0.05$ cm}
    \label{fig:sim_hm_4}
  \end{subfigure}
  \begin{subfigure}[]{.3\textwidth}
    \includegraphics[width=\textwidth]{sim/result_intersection_area_at_each_point_2a}
    \caption{Two $3$ microphone arrays are placed $1$ meter apart. The average error is $2.60$ cm}
    \label{fig:sim_hm_2_array}
  \end{subfigure}
  \caption{Error heatmap for different array configurations}
  \label{fig:sim_hm}
\end{figure*}

Each microphone pair generates a hyperbolic region, and the source location is in the intersection of these regions. The area of the intersection region is a measure of the localization accuracy: the smaller the area, the more certain we are about the source location. To evaluate an array's accuracy in a region, we can place sound source at predetermined grid points in the region and look at the intersection area for each tested point in the grid. The center location of the intersection region can be used as the localization estimate to calculate localization error. Results for a few different microphone array configurations are presented in fig~\ref{fig:sim_hm}.

Fig~\ref{fig:sim_hm_3} shows the accuracy when microphones are placed at the three vertices of a $20$ cm equilateral triangle. The region inside the array has good accuracy. However, for regions along the line connecting any two microphones, the accuracy drops significantly. The average error across the region is $18.6$ cm.

To evaluate how adding one microphone(without increasing the array size) improves accuracy, another microphone is added to the array at $(0,0)$. Result is presented in fig~\ref{fig:sim_hm_3_p1}. Addition of the new microphone only slightly improved the accuracy around the array region. The average error dropped from $18.6$ cm to $17.1$ cm. Regions near lines connecting microphones still have significantly larger uncertainty region.

To evaluate the array size's impact on accuracy, the size of the original array (as in fig~\ref{fig:sim_hm_3}) is increased by a factor of $2$. The result is presented in fig~\ref{fig:sim_hm_3_2x}. The overall uncertainty area decreased across the region. The average error improved to $10.04$ cm.
